Benchmarking

### 2023-08-12 Misc/benchmarks.qmd
brms threading benchmarks

>>>>

library(ggplot2)
library(brms)

#https://github.com/paul-buerkner/brms/blob/master/vignettes/brms_threading.Rmd

set.seed(54647)
# number of observations
N <- 1E4
# number of group levels
G <- round(N / 10)
# number of predictors
P <- 3
beta <- rnorm(P)
fake <- matrix(rnorm(N * P), ncol = P)
dimnames(fake) <- list(NULL, paste0("x", 1:P))
fake <- transform(
  as.data.frame(fake),
  theta = fake %*% beta,
  g = sample.int(G, N, replace=TRUE)
)
fake  <- merge(fake, data.frame(g = 1:G, eta = rnorm(G)), by = "g")
fake  <- transform(fake, mu = theta + eta)
fake  <- transform(fake, y = rpois(N, exp(mu)))
fake <- fake[sample.int(N, N),]
rownames(fake) <- NULL

kable(head(fake, 10), digits = 3)

# Benchmarks given model with cross-product of tuning parameters CPU
# cores, grainsize and iterations. Models are run with either static
# or non-static scheduler and inits is set by default to 0 on the
# unconstrained scale. Function returns a data-frame with the
# cross-product of the tuning parameters and as result column the
# respective runtime.
benchmark_threading <- function(model, cores = 1, grainsize = 1, iter = 100, 
                                static = FALSE) {

    winfo <- extract_warmup_info(model)
    sims  <- rstan::extract(model$fit)
    init <- list(extract_draw(sims, 1))

    scaling_model <- update(
        model, refresh = 0, 
        threads = threading(1, grainsize = grainsize[1], static = static), 
        chains = 1, iter = 2, backend = "cmdstanr"
    )

    run_benchmark <- function(cores, size, iter) {
        bench_fit <- update(
            scaling_model, warmup=0, iter = iter,
            chains = 1, seed = 1234, inits = init, refresh = 0, save_warmup=TRUE,
            threads = threading(cores, grainsize = size, static = static),
            inv_metric=winfo$inv_metric[[1]],
            step_size=winfo$step_size[[1]],
            adapt_engaged=FALSE
        )
        lf <- sum(subset(nuts_params(bench_fit, inc_warmup=TRUE), Parameter=="n_leapfrog__")$Value)
        elapsed <- sum(colSums(rstan::get_elapsed_time(bench_fit$fit)))

        c(num_leapfrog=lf, runtime=elapsed)
    }

    cases <- expand.grid(cores = cores, grainsize = grainsize, iter = iter)
    res <- with(cases, mapply(run_benchmark, cores, grainsize, iter))
    cbind(cases, as.data.frame(t(res)))
}

benchmark_reference <- function(model, iter=100, inits=0) {
    winfo <- extract_warmup_info(model)
    sims  <- rstan::extract(model$fit)
    init <- list(extract_draw(sims, 1))

    ref_model <- update(
        model, refresh = 0, 
        threads = NULL,
        chains = 1, iter = 2, backend = "cmdstanr"
    )

    run_benchmark_ref <- function(iter_bench) {
        bench_fit <- update(
            ref_model, warmup=0, iter = iter_bench,
            chains = 1, seed = 1234, inits = init, refresh = 0,
            inv_metric=winfo$inv_metric[[1]],
            step_size=winfo$step_size[[1]],
            adapt_engaged=FALSE
        )

        lf <- sum(subset(nuts_params(bench_fit, inc_warmup=TRUE), Parameter=="n_leapfrog__")$Value)
        elapsed <- sum(colSums(rstan::get_elapsed_time(bench_fit$fit)))

        c(num_leapfrog=lf, runtime=elapsed)
    }

    ref <- sapply(iter, run_benchmark_ref)
    ref <- cbind(as.data.frame(t(ref)), iter=iter)
    ref
}

extract_warmup_info <- function(bfit) {
    adapt  <- lapply(rstan::get_adaptation_info(bfit$fit), strsplit, split="\\n")
    step_size  <- lapply(adapt, function(a) as.numeric(strsplit(a[[1]][[1]], " = ")[[1]][2]))
    inv_metric <- lapply(adapt, function(a) as.numeric(strsplit(sub("^# ", "", a[[1]][[3]]), ", ")[[1]]))
    list(step_size=step_size, inv_metric=inv_metric)
}

extract_draw <- function(sims, draw) lapply(sims, abind::asub, idx=draw, dim=1, drop=FALSE)

model_poisson <- brm(
  y ~ 1 + x1 + x2 + (1 | g),
  data = fake,
  family = poisson(),
  iter = 500, # short sampling to speedup example
  chains = 2,
  prior = prior(normal(0,1), class = b) +
    prior(constant(1), class = sd, group = g),
  backend = "cmdstanr",
  threads = threading(4)
)

chunking_bench <- transform(
    data.frame(chunks = 4^(0:3)),
    grainsize = ceiling(N / chunks)
)

iter_test <- c(10, 20, 40)  # very short test runs
scaling_chunking <- benchmark_threading(
  model_poisson,
  cores = 1,                         
  grainsize = chunking_bench$grainsize,  # test various grainsizes
  iter = iter_test,  
  static = TRUE  # with static partitioner
)

# run as reference the model *without* reduce_sum
ref <- benchmark_reference(model_poisson, iter_test)

https://cran.r-project.org/web/packages/brms/vignettes/brms_threading.html

https://discourse.mc-stan.org/t/using-the-apple-m1-gpus-question-from-a-noob/23089/39?page=2

```{r}
#| warning: false

pacman::p_load(tidyverse,tidybayes,brms,broom,broom.mixed,lme4,here,knitr,gt,gghalves,patchwork,ggdist,microbenchmark)
e1 <- readRDS(here("data/e1_08-04-23.rds"))
test <- e1 |> filter(expMode2 == "Test")

options(mc.cores = 4, brms.backend = "cmdstanr")

n_cores=4
bayes_seed <- 1234
n_iter=20000
n_threads=2

gprior<- c( prior(normal(800, 100), class = Intercept),
    prior(normal(400, 10), class = sigma)
  )

##############

tf3 <- system.time(
fit_int_norm <- brm(vx ~ 1 + condit, 
  data = test,
  family = gaussian(),
  iter=n_iter,
  prior = gprior,
  silent=2,
  cores=n_cores,
  threads = threading(n_threads),
  seed=bayes_seed
)
)
tf3

tf4 <- system.time(
fit_int_norm <- brm(vx ~ 1 + condit +(1|bandInt) + (1|id),
  data = test |> filter(id %in% 1:15),
  family = gaussian(),
  iter=n_iter,
  prior = gprior,
  silent=2,
  cores=n_cores,
  threads = threading(n_threads),
  seed=bayes_seed
)
)
tf4

#########

tf <- system.time(
fit_int_norm <- brm(vx ~ 1 + condit, 
  data = test,
  family = gaussian(),
  iter=n_iter,
  prior = gprior,
  silent=2,
  cores=n_cores,
  seed=bayes_seed
)
)
tf

tf2 <- system.time(
fit_int_norm <- brm(vx ~ 1 + condit +(1|bandInt) + (1|id),
  data = test |> filter(id %in% 1:15),
  family = gaussian(),
  iter=n_iter,
  prior = gprior,
  silent=2,
  cores=n_cores,
  seed=bayes_seed
)
)
tf2

##########

cat(paste0("int only gaussian = ",round(tf["elapsed"],2)," \n ",
          # "foreach parallel time=",round(tfP["elapsed"],2)," \n ",
           "irt version = ",round(tf2["elapsed"],2), " \n ",
           "int only threading = ",round(tf3["elapsed"],2)," \n ",
          "irt  threading = ",round(tf4["elapsed"],2)," \n "

          ))

```
### default brms settings (2K iterations)
int only gaussian = 6.85 
irt version = 12.15

int only gaussian = 7.51 
irt version = 12

int only gaussian = 6.67 
irt version = 11.82 
int only threading = 7.49 
irt  threading = 11.33

### default brms settings (4K iterations) - 2 threads
 int only gaussian = 8.1 
irt version = 14.69 
int only threading = 6.99 
irt  threading = 15.24

### default brms settings (4K iterations) - 4 threads
brms int only gaussian = 8.11 
irt version = 15.05 
int only threading = 7.28 
irt  threading = 16.95

### default brms settings (4K iterations) - 4 threads - 2 cores
int only gaussian = 11.27 
irt version = 21.72 
int only threading = 8.39 
irt  threading = 22.67

### default brms settings (4K iterations) - 2 threads - 2 cores
int only gaussian = 11.27 
irt version = 21.81 
int only threading = 9.38 
irt  threading = 20.63

### default brms settings (4K iterations) - 8 threads - 4 cores
int only gaussian = 8.27 
irt version = 15.4 
int only threading = 7.69 
irt  threading = 21.28

int only gaussian = 8.37 
irt version = 15.18 
int only threading = 8.02 
irt  threading = 21.47

### default brms settings (4K iterations) - 8 threads - 8 cores

int only gaussian = 8.11 
irt version = 14.88 
int only threading = 7.92 
irt  threading = 20.72

### default brms settings (10K iterations) - 8 threads - 8 cores

int only gaussian = 11.38 
irt version = 23.64 
int only threading = 11.58 
irt  threading = 37.1

int only gaussian = 11.53 
irt version = 23.76 
int only threading = 12 
irt  threading = 37.45

### default brms settings (10K iterations) - 8 threads - 4 cores

int only gaussian = 11.48 
irt version = 23.29 
int only threading = 10.91 
irt  threading = 36.54

### default brms settings (10K iterations) - 4 threads - 4 cores

int only gaussian = 11.99 
irt version = 24.25 
int only threading = 10.13 
irt  threading = 29.13

### default brms settings (10K iterations) - 4 threads - 1 cores

int only gaussian = 29.35 
irt version = 66.61 
int only threading = 18.97 
irt  threading = 71.21

### default brms settings (10K iterations) - 2 threads - 2 cores

int only gaussian = 17.98 
irt version = 38.31 
int only threading = 13.87 
irt  threading = 37.78

### default brms settings (10K iterations) - 1 threads - 4 cores

int only gaussian = 11.14 
irt version = 22.89 
int only threading = 12.25 
irt  threading = 27.03

int only gaussian = 11.35 
 irt version = 22.99 
 int only threading = 12.28 
 irt  threading = 27.38

### default brms settings (10K iterations) - 2 threads - 4 cores

int only gaussian = 11.22 
irt version = 23.07 
int only threading = 9 
irt  threading = 22.34

int only gaussian = 11.31 
 irt version = 23.14 
 int only threading = 9.49 
 irt  threading = 22.21

### default brms settings (10K iterations) - 3 threads - 4 cores

int only gaussian = 11.15 
irt version = 22.88 
int only threading = 9.4 
irt  threading = 26.75

int only gaussian = 11 
 irt version = 22.88 
 int only threading = 9.5 
 irt  threading = 25.7

### (20K iterations) - 1 threads - 4 cores

int only gaussian = 18.24 
 irt version = 40.37 
 int only threading = 20.32 
 irt  threading = 48.94

int only gaussian = 17.8 
 irt version = 39.26 
 int only threading = 20.54 
 irt  threading = 48.14

 int only gaussian = 18.12 
 irt version = 40.31  
 int only threading = 20.6 
 irt  threading = 40.02 - when I commented out the threading specification

### (20K iterations) - 2 threads - 4 cores

int only gaussian = 17.82 
irt version = 39.34 
int only threading = 14.22 
irt  threading = 37.73
